---
title: "Sentiment analysis"
mainfont: Arial
fontsize: 12pt
urlcolor: blue
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 1
    toc_float: true
    df_print: paged
    theme: paper
    pandoc_args: --output=Impractical_4.html
---

# Introduction

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

Welcome to the fourth practical of the course "Introduction to Text Mining with R". In this practical, we will analyze sentiments in a text data set, first by using dictionaries and then we will classify sentiment in text reviews comparing a dictionary-based method and a neural network model.

In this practical we are going to use the following packages:

```{r load_packages, message = FALSE, warning = FALSE}
library(tm)
library(tidytext)
library(dplyr)

```

---

# Dictionary-based sentiment analysis

---

---

## Read data

---

1. **Read the "taylor_swift.csv" dataset from the data folder.**

---

---

2. **First we must preprocess the corpus. Create a document-term matrix from the `lyrics` column of the `ts` data frame. Complete the following preprocessing steps:**
- convert to lower
- remove stop words
- remove numbers
- remove punctuation.

---

---

3. **Inspect the dtm object and convert it to a dataframe.**

---

---

## Sentiment dictionaries

---

4. **We're going to use sentiment dictionaries from the `tidytext` package. Using the `get_sentiments` function, load the "bing" and "afinn" dictionaries and store them in two objects called `bing_sentiments` and `afinn_sentiments`. **

The tidytext package contains 4 general purpose lexicons in the sentiments dataset.

AFINN - listing of english words rated for valence between -5 and +5
bing - listing of positive and negative sentiment
nrc - list of English words and their associations with 8 emotions (anger, fear, anticipation, trust, surprise, sadness, joy, and disgust) and 2 sentiments (negative and positive); binary
loughran - list of sentiment words for accounting and finance by category (Negative, Positive, Uncertainty, Litigious, Strong Modal, Weak Modal, Constraining)

---

---

5. **The afinn_sentiments has the rating for valence between -5 and +5, while the bing_sentiments contains listing of positive and negative sentiment. Add a column to `bing_sentiments` called `score`. This column should hold a "1" for positive words and "-1" for negative words.**

---

---

---

## Calculate sentiment score for each lyric

---

6. **Create a dataframe that holds all the words in the dtm object along with their sentiment score.**

---

---

7. **To calculate a score for each lyric, multiply your dtm object by the scoring dataframe.**

---

---

8. **Add the calculated scores for each lyric to the original `taylor_swift_lyrics` dataframe.**

---

---

9. **Using the code you wrote above, we made a function that gets 1) a vector of texts, and 2) a sentiment dictionary (i.e. a data frame with words and scores), and returns a vector of sentiment scores for each text. Use this function to repeat your analysis with the `afinn` sentiment dictionary.**

---

10. **Compare the bing and afinn dictionaries by finding out which the most and least positive Taylor Swift album is.**

---

---

# Sentiment analysis of product reviews

---

11. **Load the final `computer_531` dataframe from the first practical and  to read data from the 'computer.txt' file.**

---

---

12. **Apply the `sentiment_score` function on the reviews in the dataframe with both the bing and afinn dictionaries.**

---

---

13. **Create a confusion matrix. Accuracy, precision, recall and F1 measures.**

---

---

14. **From the `neuralnet` package, we want to apply a multi-layer perceptron to this dataset and train a model for prediction of the sentiments.**

---


---

15. **Compare nn1, bing and affine ...**

---

---

16. **Visualise the results**

---


---

# Summary

---



---

End of Practical

---
