---
title: "Text Clustering and Topic Modeling"
mainfont: Arial
fontsize: 12pt
urlcolor: blue
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 1
    toc_float: true
    df_print: paged
    theme: paper
    pandoc_args: --output=Practical_5.html
---

# Introduction

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

Welcome to the fifth practical of the course "Introduction to Text Mining with R". In this practical, we will apply a clustering and a topic modeling approach on news articles and cluster them into different categories. 

In this practical we are going to use the following packages:

```{r load_packages, message = FALSE, warning = FALSE}
library(tm)
library(tidytext)
library(dplyr)

```

---

# 

---

---

## Read data

---

1. **The dataset that we are going to use is the BBC dataset from practical 3. The BBC dataset consists of 2225 documents and 5 categories: business, entertainment, politics, sport, and tech. For text clustering and topic modeling, we will ignore the labels but we will use them while evaluating models. Load the Rda file of the news dataset.**

---

2. **First we must preprocess the corpus. Create a document-term matrix from the `news` column of the dataframe. Complete the following preprocessing steps:**
- convert to lower
- remove stop words
- remove numbers
- remove punctuation
- use tfidf weighting
- convert dtm to a dataframe

---

3. **Now we can implement k-means clustering, starting out with three centers. What does the output look like?**

---

## Clustering methods
## Choosing k

---

4. ** **

We used k = 3 but how do we know that’s right? There are lots of complicated or “more art than science” ways of choosing k. One way is to look at the total within-cluster sum of squares and see if it stops dropping off so quickly at some value for k. We can get that from another verb from broom, glance(); let’s try lots of values for k and see what happens to the total sum of squares.

---

5. **Fit k-means again with the optimal k.**

---


---

16. **Visualise the results**

---

```{r 16, include = TRUE}

```


---



---

# Summary

---



---

End of Practical

---
