---
title: "Practical 6: Word embedding approaches"
author: "Ayoub Bagheri"
date: "Introduction to Text Mining with R"
mainfont: Arial
fontsize: 12pt
urlcolor: blue
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 1
    toc_float: true
    df_print: paged
    theme: paper
    pandoc_args: --output=Practical_6.html
---

# Introduction

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

Welcome to the last practical of the course "Introduction to Text Mining with R". In this practical, we will apply word embedding approaches. 

In this practical, we are going to use the following packages:

```{r load_packages, message = FALSE, warning = FALSE}
library(text2vec)
library(tidyverse)
library(tidytext)
```

---

# Word embeddings

---

A key idea in working with text data concerns representing words as numeric quantities. There are a number of ways to go about this, and we have actually already done so. An additional method that we want to explore today is word embeddings. Word embedding techniques such as word2vec and GloVe use neural networks approaches to construct word vectors. With these vector representations of words, we can see how similar they are to each other, and perform other tasks based on that information. Here are two famous examples:

king - man + woman = queen


Paris - France + Germany = Berlin


The first part of the practical leverages the data provided in the harrypotter package. This package supplies the first seven novels in the Harry Potter series. You can load the harrypotter package with the following:

```{r}
#devtools::install_github("bradleyboehmke/harrypotter")
library(harrypotter) # Not to be confused with the CRAN palettes package
```

The aim of [devtools](https://www.r-project.org/nosvn/pandoc/devtools.html) is to make your life easier by providing R functions that simplify many common tasks.

---

1. **Use the code below to load the first seven novels in the Harry Potter series. View the data sets.**

---

```{r 1, message = FALSE, warning = FALSE, include = TRUE}
hp_books <- c("philosophers_stone", "chamber_of_secrets",
              "prisoner_of_azkaban", "goblet_of_fire",
              "order_of_the_phoenix", "half_blood_prince",
              "deathly_hallows")

hp_words <- list(
  philosophers_stone,
  chamber_of_secrets,
  prisoner_of_azkaban,
  goblet_of_fire,
  order_of_the_phoenix,
  half_blood_prince,
  deathly_hallows
) %>%
  # name each list element
  set_names(hp_books) %>%
  # convert each book to a data frame and merge into a single data frame
  map_df(as_tibble, .id = "book") %>%
  # convert book to a factor
  mutate(book = factor(book, levels = hp_books)) %>%
  # remove empty chapters
  filter(!is.na(value)) %>%
  # create a chapter id column
  group_by(book) %>%
  mutate(chapter = row_number(book))

head(hp_words)

```

---

2. **Convert the hp_words object into a dataframe and use the unnest_tokens() function from the tidytext package to tokenize the data frame.**

---
```{r 2, message = FALSE, warning = FALSE, include = TRUE}


```

---

3. **Remove the stop words from the tokenized data frame.**

---
```{r 3, message = FALSE, warning = FALSE, include = TRUE}


```

---

4. **Create a vocabulary of unique terms using the create_vocabulary() function from the text2vec package and remove the words that appear less than 5 times.**

---
```{r 4, message = FALSE, warning = FALSE, include = TRUE}


```

---

5. **The next step is to create a token co-occurrence matrix (TCM). The definition of whether two words occur together is arbitrary. First create a vocab_vectorizer, then, use a window of 5 for context words to create the TCM.**

---

```{r 5, message = FALSE, warning = FALSE, include = TRUE}


```

---

6. **Use the GlobalVectors to fit the word vectors on our data set. Choose the embedding size (rank) equal to 50, and the maximum number of co-occurrences to use in the weighting function equal to 10. Train word vectors in 50 iterations if you have the (memory / cpu) resources. Also check the other input arguments of the fit_transform function [here](https://www.rdocumentation.org/packages/text2vec/versions/0.5.1/topics/GlobalVectors).**

---

```{r 6, message = FALSE, warning = FALSE, include = TRUE}


```

---

7. **Extract the word vectors and save the sum of them for further questions.**

---

```{r 7, message = FALSE, warning = FALSE, include = TRUE}


```

---

8. **Find the most similar words to words "harry", "death", and "love". Use the cosine similarity measure with the function sim2. Set the norm argument to "l2".**

---

```{r 8, message = FALSE, warning = FALSE, include = TRUE}


```

---

9. **Add the word vector of "harry" with the word vector of "love" and subtract them from the word vector of "death". What are the top terms in your result?**

---

```{r 9, message = FALSE, warning = FALSE, include = TRUE}


```

---

# Wikipedia word embeddings

---

10. **Here we want to repeat the same analysis as for Harry Potter novel series with texts from Wikipedia. Start with the code below and train the word vectors using the wiki object.**

---

```{r}
text8_file <- "data/text8/text8"
if (!file.exists(text8_file)) {
  download.file("http://mattmahoney.net/dc/text8.zip", "data/text8.zip")
  unzip("data/text8.zip", files = "text8", exdir = "data/texts_raw/")
}
wiki <- readLines(text8_file, n = 1, warn = FALSE)

```


```{r 10, message = FALSE, warning = FALSE, include = TRUE}


```

---

11. **Use the Wikipedia word embeddings and try the two famous examples below.**

king - man + woman = queen


Paris - France + Germany = Berlin

---

```{r 11, message = FALSE, warning = FALSE, include = TRUE}


```
---

# TensorFlow and Keras

In this part of the practical, we will show an example of loading pre-trained word vectors and fine-tune them for our purpose of sentiment classification for movie reviews. First, we need to install the TensorFlow and Keras packages for R.

The tensorflow package provides code completion and inline help for the [TensorFlow API](https://www.tensorflow.org/api_docs/python/tf/all_symbols) when running within the RStudio IDE. The TensorFlow API is composed of a set of Python modules that enable constructing and executing TensorFlow graphs.

To get started, we need to use the devtools package from CRAN. If you do not have it install it first:

```{r}
# This line is commented because I already have the package installed!
# install.packages("devtools")

```

The aim of devtools is to make package development easier by providing R functions that simplify and expedite common tasks.

Then, install the tensorflow R package from GitHub as follows:

```{r}
# devtools::install_github("rstudio/tensorflow")

```

Then, use the install_tensorflow() function to install TensorFlow:

```{r}
# tensorflow::install_tensorflow(package_url = "https://files.pythonhosted.org/packages/66/8e/19d47c7b9e2629c37fb3b2febb09a45d4ee7055e6bcb7ff790d8604d822c/tensorflow-2.9.1-cp310-cp310-win_amd64.whl")
# tensorflow::install_tensorflow(version = "2.9.1", method = "conda", envname = "r-reticulate") # in case line 312 doesn't work

library(tensorflow) 

```

The provided url just installs the latest tensorflow version, you can also run this line without providing any argument to the install_tensorflow function.

Finally, you can confirm that the installation succeeded with:

```{r}
tmr <- tf$constant("Text Mining with R!")
print(tmr)

```

This will provide you with a default installation of TensorFlow suitable for getting started with the tensorflow R package. See the article on installation (https://tensorflow.rstudio.com/installation/) to learn about more advanced options, including installing a version of TensorFlow that takes advantage of Nvidia GPUs if you have the correct CUDA libraries installed.


To install the Keras package you first run either of the following lines:
```{r}
# install.packages("keras")

```

```{r}
# devtools::install_github("rstudio/keras")

```

Then, use the install_keras() function to install keras:

```{r}
# install_keras()

```

The Keras R interface uses the TensorFlow backend engine by default. This will provide you with default CPU-based installations of Keras and TensorFlow. If you want a more customized installation, e.g. if you want to take advantage of NVIDIA GPUs, see the documentation for install_keras() and the article on installation (https://tensorflow.rstudio.com/installation/).


# Sentiment classification with pre-trained word vectors

Now we have TensorFlow and Keras ready for fine-tuning of pre-trained word embeddings for sentiment classification of movie reviews. 

Remember to load the following libraries:
```{r}
library(keras)
```

---

12. **For this purpose, we want to use [GloVe](https://nlp.stanford.edu/projects/glove/) pretrained word vectors. In the data folder you find these word vectors which were trained on Wikipedia 2014 and Gigaword Fifth Edition containing 6B tokens, 400K vocab, uncased, 50d, 100d, 200d, & 300d vectors. You can also manually download them or use the code below for this purpose.**

---

```{r 12, message = FALSE, warning = FALSE, include = TRUE}
# Download Glove vectors if necessary and save them in your data folder
# if (!file.exists('glove.6B.zip')) {
#   download.file('https://nlp.stanford.edu/data/glove.6B.zip',destfile = 'data/glove.6B.zip')
#   unzip('glove.6B.zip')
# }

```

---

13. **Load the pre-trained word vectors from the file 'glove.6B.300d.txt' (if you have memory issues load the file 'glove.6B.50d.txt' instead).**

---

```{r 13, message = FALSE, warning = FALSE, include = TRUE}


```

---

14. **IMDB movie reviews is a labeled data set available in the text2vec package. This data set consists of 5000 IMDB movie reviews, specially selected for sentiment analysis. The sentiment of the reviews is binary, meaning an IMDB rating < 5 results in a sentiment score of 0, and a rating >=7 has a sentiment score of 1. No individual movie has more than 30 reviews. Load this data set and convert that to a dataframe.**

---

```{r 14, message = FALSE, warning = FALSE, include = TRUE}


```

---

15. **Define the parameters of your keras model with a maximum of 10000 words, maxlen of 60 and word embedding size of 300 (if you had memory problems change the embedding dimension to 50).**

---

```{r 15, message = FALSE, warning = FALSE, include = TRUE}

```

---

16. **Use the text_tokenizer function from keras and tokenize the imdb review data using a maximum of 10000 words.**

---

```{r 16, message = FALSE, warning = FALSE, include = TRUE}


```

---

17. **Transform each text in `movie_review$review` to a sequence of integers and get indices instead of words, later pad the sequence.**

---

```{r 17, message = FALSE, warning = FALSE, include = TRUE}


```

---

18. **Convert the word indices into a dataframe.**

---

```{r 18, message = FALSE, warning = FALSE, include = TRUE}


```

---

19. **Join the dataframe of indices of words from the IMDB reviews with GloVe pre-trained word vectors.**

---

```{r 19, message = FALSE, warning = FALSE, include = TRUE}


```

---

20. **Extract the outcome variable from the sentiment column in the original dataframe and name it y_train.**

---

```{r 20, message = FALSE, warning = FALSE, include = TRUE}


```

---

21. **Use the Keras functional API and create a neural network model as below. Can describe this model?**

---

```{r 21, message = FALSE, warning = FALSE, include = TRUE}


```

---

22. **Compile the model with an 'adam' optimizer, and the binary_crossentropy loss. You can choose accuracy or AUC for the metrics.**

---

```{r 22, message = FALSE, warning = FALSE, include = TRUE}


```

---

23. **Fit the model with  5 or 10 epochs (iterations), batch_size = 32, and validation_split = 0.2. Check the training performance versus the validation performance.**

---

```{r 23, message = FALSE, warning = FALSE, include = TRUE}


```

```{r}


```


---

# Summary

---

In this practical, we learned about:

- Word embeddings
- Pre-trained word vectors
- Text2vec, Keras, TensorFlow
- More on [Keras Training Visualization](https://keras.rstudio.com/articles/training_visualization.html)

---

End of Practical

---
